Response to Referees

### 2025-06-10 ms/ms.Rmd
address RC2.8: concordance between different VI methods

We calculated variable importance with the *vip* R package (v0.4), by three different methods: *FIRM*, *permutation*, and *Shapley* [@greenwellVariableImportancePlots2020].
*FIRM* values measure the flatness of the partial dependence plot, *permutation* values measure the decrease in model performance when the predictor is permuted, and *Shapley* values are aggregated from local, game-theoretical measures of variable importance [@greenwellVariableImportancePlots2020]. 
*Permutation* values were obtained from ten iterations, with root mean square error as the performance measure.

@@

@@

At both sites, elevation and Multi-Resolution Valley Bottom Flatness were important predictors (Fig. \@ref(fig:varImp)).
At Skrimfjella these two predictors were of similar importance, while at Ørskogfjellet elevation was more important than Multi-Resolution Valley Bottom Flatness.
DMK was also important -- the shallow class in particular -- but only at Ørskogfjellet.
Some realizations of the hydrological predictors Topographic Wetness Index and Depth-to-Water showed considerable importance, while others showed little -- with no consistency between sites.
For example, TWI5m, DTW40000, and DTW2500 rounded out the top five most important variables at Skrimfjella, while TWI20m and TWI50m were least important, other than DMK.
At both sites, the most important realizations of hydrological predictors were more important than the simple terrain indices slope, Terrain Ruggedness Index, Topographic Position Index, and roughness.
The radiometric predictor uranium ground concentration showed moderate importance at Skrimfjella.

```{r varImp, echo = FALSE, out.width="100%", fig.cap = "Global variable importance in the best-performing models at Skrimfjella (a) and Ørskogfjellet (b), as measured by three different metrics: Shapley values, permutation importance, and the Feature Importance Ranking Measure [@greenwellVariableImportancePlots2020]. Variables removed due to collinearity are shown to the right of the variable with which they are most correlated."}
knitr::include_graphics("figures/variable_importance.pdf")
```

>>>>

We calculated variable importance with the *vip* R package (v0.4), by three different methods: *FIRM*, *permutation*, and *Shapley* [@greenwellVariableImportancePlots2020].
*FIRM* values measure the flatness of the partial dependence plot, *permutation* values measure the decrease in model performance when the predictor is permuted, and *Shapley* values are aggregated from local, game-theoretical measures of variable importance [@greenwellVariableImportancePlots2020].
Since *FIRM* reflects the flatness of the partial dependence plot, it captures functional complexity rather than overall predictive impact.
*Permutation* values were obtained from ten iterations, with root mean square error as the performance measure.

@@

The *permutation* and *Shapley* methods of variable importance were highly correlated at both sites, while the *FIRM* method was less correlated with the other two (Fig. \@ref(fig:varImp)).

@@

At both sites, elevation and Multi-Resolution Valley Bottom Flatness were important predictors (Fig. \@ref(fig:varImp)).
At Skrimfjella these two predictors were of similar importance, while at Ørskogfjellet elevation had higher predictive impact (*permutation* and *Shapley*) but lower functional complexity (*FIRM*) than Multi-Resolution Valley Bottom Flatness.
DMK was also important -- the shallow class in particular -- but only at Ørskogfjellet.
Some realizations of the hydrological predictors Topographic Wetness Index and Depth-to-Water showed considerable importance, while others showed little -- with no consistency between sites.
For example, TWI5m, DTW40000, and DTW2500 rounded out the top five most important variables at Skrimfjella, while TWI20m and TWI50m were least important, other than DMK.
At both sites, the most important realizations of hydrological predictors were more important than the simple terrain indices slope, Terrain Ruggedness Index, Topographic Position Index, and roughness.
The radiometric predictor uranium ground concentration showed moderate importance at Skrimfjella.

```{r varImp, echo = FALSE, out.width="100%", fig.cap = "Global variable importance in the best-performing models at Skrimfjella (a) and Ørskogfjellet (b), as measured by three different metrics: Shapley values, permutation importance, and the Feature Importance Ranking Measure \\citep{greenwellVariableImportancePlots2020}. Variables removed due to collinearity are shown to the right of the variable with which they are most correlated. Spearman rank correlations between the three metrics are displayed in matrices."}
knitr::include_graphics("figures/variable_importance.pdf")
```

### 2025-06-10 ms/ms.tex
address RC2.8: concordance between different VI methods

We calculated variable importance with the \emph{vip} R package (v0.4), by three different methods: \emph{FIRM}, \emph{permutation}, and \emph{Shapley} \citep{greenwellVariableImportancePlots2020}.
\emph{FIRM} values measure the flatness of the partial dependence plot, \emph{permutation} values measure the decrease in model performance when the predictor is permuted, and \emph{Shapley} values are aggregated from local, game-theoretical measures of variable importance \citep{greenwellVariableImportancePlots2020}.
\emph{Permutation} values were obtained from ten iterations, with root mean square error as the performance measure.

@@

@@

At both sites, elevation and Multi-Resolution Valley Bottom Flatness were important predictors (Fig. \ref{fig:varImp}).
At Skrimfjella these two predictors were of similar importance, while at Ørskogfjellet elevation was more important than Multi-Resolution Valley Bottom Flatness.
DMK was also important -- the shallow class in particular -- but only at Ørskogfjellet.
Some realizations of the hydrological predictors Topographic Wetness Index and Depth-to-Water showed considerable importance, while others showed little -- with no consistency between sites.
For example, TWI5m, DTW40000, and DTW2500 rounded out the top five most important variables at Skrimfjella, while TWI20m and TWI50m were least important, other than DMK.
At both sites, the most important realizations of hydrological predictors were more important than the simple terrain indices slope, Terrain Ruggedness Index, Topographic Position Index, and roughness.
The radiometric predictor uranium ground concentration showed moderate importance at Skrimfjella.

\begin{figure}
\includegraphics[width=1\linewidth]{figures/variable_importance} \caption{Global variable importance in the best-performing models at Skrimfjella (a) and Ørskogfjellet (b), as measured by three different metrics: Shapley values, permutation importance, and the Feature Importance Ranking Measure [@greenwellVariableImportancePlots2020]. Variables removed due to collinearity are shown to the right of the variable with which they are most correlated.}\label{fig:varImp}
\end{figure}

>>>>

We calculated variable importance with the \emph{vip} R package (v0.4), by three different methods: \emph{FIRM}, \emph{permutation}, and \emph{Shapley} \citep{greenwellVariableImportancePlots2020}.
\emph{FIRM} values measure the flatness of the partial dependence plot, \emph{permutation} values measure the decrease in model performance when the predictor is permuted, and \emph{Shapley} values are aggregated from local, game-theoretical measures of variable importance \citep{greenwellVariableImportancePlots2020}.
Since \emph{FIRM} reflects the flatness of the partial dependence plot, it captures functional complexity rather than overall predictive impact.
\emph{Permutation} values were obtained from ten iterations, with root mean square error as the performance measure.

@@

The \emph{permutation} and \emph{Shapley} methods of variable importance were highly correlated at both sites, while the \emph{FIRM} method was less correlated with the other two (Fig. \ref{fig:varImp}).

@@

At both sites, elevation and Multi-Resolution Valley Bottom Flatness were important predictors (Fig. \ref{fig:varImp}).
At Skrimfjella these two predictors were of similar importance, while at Ørskogfjellet elevation had higher predictive impact (\emph{permutation} and \emph{Shapley}) but lower functional complexity (\emph{FIRM}) than Multi-Resolution Valley Bottom Flatness.
DMK was also important -- the shallow class in particular -- but only at Ørskogfjellet.
Some realizations of the hydrological predictors Topographic Wetness Index and Depth-to-Water showed considerable importance, while others showed little -- with no consistency between sites.
For example, TWI5m, DTW40000, and DTW2500 rounded out the top five most important variables at Skrimfjella, while TWI20m and TWI50m were least important, other than DMK.
At both sites, the most important realizations of hydrological predictors were more important than the simple terrain indices slope, Terrain Ruggedness Index, Topographic Position Index, and roughness.
The radiometric predictor uranium ground concentration showed moderate importance at Skrimfjella.

\begin{figure}
\includegraphics[width=1\linewidth]{figures/variable_importance} \caption{Global variable importance in the best-performing models at Skrimfjella (a) and Ørskogfjellet (b), as measured by three different metrics: Shapley values, permutation importance, and the Feature Importance Ranking Measure \citep{greenwellVariableImportancePlots2020}. Variables removed due to collinearity are shown to the right of the variable with which they are most correlated. Spearman rank correlations between the three metrics are displayed in matrices.}\label{fig:varImp}
\end{figure}

### 2025-06-10 ms/submission/AuthorComments-egusphere-2025-1046.qmd
address RC2.8: concordance between different VI methods

>>>>

Thank you for this suggestion. We have added text to the _Materials and methods_ and _Results_ sections to clarify how the three variable importance methods relate to each other. We have also added rank correlations to our figure on variable importance, to highlight the similarities and differences. In short, we find that Shapley and permutation importance are very similar for our application, while FIRM values capture a different aspect of variable importance.

Since the different methods are broadly congruent and their differences do not affect our interpretation much -- and in the interest of brevity -- we do not go into this point in our _Discussion_.

We calculated variable importance with the *vip* R package (v0.4), by three different methods: *FIRM*, *permutation*, and *Shapley* [@greenwellVariableImportancePlots2020].
*FIRM* values measure the flatness of the partial dependence plot, *permutation* values measure the decrease in model performance when the predictor is permuted, and *Shapley* values are aggregated from local, game-theoretical measures of variable importance [@greenwellVariableImportancePlots2020].
[Since *FIRM* reflects the flatness of the partial dependence plot, it captures functional complexity rather than overall predictive impact.]{.added}
*Permutation* values were obtained from ten iterations, with root mean square error as the performance measure.

[The *permutation* and *Shapley* methods of variable importance were highly correlated at both sites, while the *FIRM* method was less correlated with the other two (Fig. \@ref(fig:varImp)).]{.added}

At both sites, elevation and Multi-Resolution Valley Bottom Flatness were important predictors (Fig. \@ref(fig:varImp)).
At Skrimfjella these two predictors were of similar importance, while at Ørskogfjellet elevation [was more important than Multi-Resolution Valley Bottom Flatness]{.deleted} [had higher predictive impact (*permutation* and *Shapley*) but lower functional complexity (*FIRM*) than Multi-Resolution Valley Bottom Flatness]{.added}.
:::
